---
output: reprex::reprex_document
knit: reprex::reprex_render
---

Testing out the fitting using nimble
Louis Freeland-Haynes
Date: 2024/02/27
This script contains the model from
   Rhinehart, Tessa A.; Turek, Daniel; Kitzes, Justin. 2022.
   A continuous-score occupancy modeling framework for incorporating
   uncertain machine learning output in autonomous biodiversity surveys.
   Methods in Ecology and Evolution.

 Note that the indices for mu and sigma are different in this file (mu[1], mu[2])
 compared to the manuscript (mu_0, mu_1) because of R's indexing.


```{r}
# install all the libraries
install.packages("nimble")
install.packages("tidyverse")
install.packages("rlist")
install.packages("bayesplot")
install.packages("readr")
```

```{r}
library(nimble)
library(tidyverse) # I think tidyverse required
library(rlist)
library(bayesplot)
library(readr)
library(ggplot2)
library(dplyr)    

```

```{r}

# read in a csv file from simulated_data directory 
# set the working directory to the location of the data
setwd("/media/auk/projects/lfh/MIPR_9_Extension_tessa/")

# read the csv, the first row is the header
data <- read_csv("./simulated_data/simulated_data_lambda_0.05_p_call_0.01.csv")
# this is a tibble where each row is a site and each column is a file's score
# turn these into rlists
scores <- as_rlist(data)
```

```{r}
# find the max score
max_score <- max(scores)
# print this
print(max_score)
```


```{r}
library(nimble)
library(tidyverse) # I think tidyverse required
library(rlist)
library(bayesplot)
####### CONTINUOUS-SCORE MODEL, NO COVARIATES #######

## NIMBLE custom distribution for score:
dContinuousScore <- nimbleFunction(
  run = function(x = double(), theta = double(), mu = double(1), sigma = double(1), z = double(), annotation = double(), log = integer()) {
    if (is.na(annotation)) {
      lp <- (1 - z) * dnorm(x, mu[1], sigma[1], log = TRUE) +
        z * log((1 - theta) * dnorm(x, mu[1], sigma[1]) + theta * dnorm(x, mu[2], sigma[2]))
    } else if (annotation == 0) {
      lp <- (1 - z) * dnorm(x, mu[1], sigma[1], log = TRUE) +
        z * (log(1 - theta) + dnorm(x, mu[1], sigma[1], log = TRUE))
    } else if (annotation == 1) {
      if (z != 1) stop("error with z value")
      lp <- log(theta) + dnorm(x, mu[2], sigma[2], log = TRUE)
    } else {
      stop("unknown value of annotation")
    }
    returnType(double())
    return(lp)
  }
)

rContinuousScore <- nimbleFunction(
  run = function(n = double(), theta = double(), mu = double(1), sigma = double(1), z = double(), annotation = double()) {
    print("should never call rContinuousScore function")
    returnType(double())
    return(1)
  }
)

registerDistributions(list(
  dContinuousScore = list(
    BUGSdist = "dContinuousScore(theta, mu, sigma, z, annotation)",
    types = c("mu = double(1)", "sigma = double(1)")
  )
))


## NIMBLE model code and model object:
code <- nimbleCode({
  lambda ~ dunif(0, 10) # overall mean in abundance across all sites (site Ns are drawn from poisson(lambda)
  p_call ~ dunif(0, 1) # probability bird calls in any one time window

  mu[1] ~ dunif(-15, 0)
  mu[2] ~ dunif(0, 15)

  for (i in 1:2) {
    sigma[i] ~ dunif(0.1, 10)
  }

  for (i in 1:NSITES) {
    abundance[i] ~ dpois(lambda) ### JAN20-2023
    z[i] <- step(abundance[i] - 1) # step(x) = 1 at x>=0, so need to subtract 1 from it

    p_at_least_1_singer[i] <- (1 - (1 - p_call)**abundance[i]) # probability any 1 period contains no singers
    num_positives[i] ~ dbinom(prob = p_at_least_1_singer[i], NFILES) # num positives
    theta[i] <- num_positives[i] / NFILES

    for (j in 1:NFILES) {
      score[i, j] ~ dContinuousScore(theta = theta[i], mu = mu[1:2], sigma = sigma[1:2], z = z[i], annotation = annotation[i, j])
    }
  }
  constraint ~ dconstraint(mu[2] >= mu[1]) # Assumption to make the two classifier distributions identifiable
})

```
```{r}

############## Fitting the data ##########

# Set working directory
# If you are using RStudio and this gives you an error, Source the code instead of Running it
this.dir <- dirname(parent.frame(2)$ofile)
setwd(this.dir)

# Set up data and results save directories
data_save_path <- "/media/auk/projects/lfh/MIPR_9_Extension_tessa/simulated_data/"
results_save_path <- "/media/auk/projects/lfh/MIPR_9_Extension_tessa/mcmc_results/"

nimbleOptions(clearNimbleFunctionsAfterCompiling = TRUE)

# Read in data
for (real_lambda in seq(0, 4, 0.2)) {
  data_file <- sprintf("%ssimulated_scores%g.Rdata", data_save_path, real_lambda)
  load(data_file)

  # Set up the constants, data, and initial value vectors needed by NIMBLE model
  constants <- list(
    NSITES = NSITES,
    NFILES = NFILES
  )
  data <- list(
    z = rep(c(NA), NSITES),
    annotation = annotation_all,
    score = scores,
    constraint = 1 # Enforces constraint
  )

  inits <- list(
    mu = c(NA, NA),
    sigma = c(NA, NA),
    z = rep(c(NA), NSITES),
    lambda = NA,
    p_call = NA
  )

  # Create the NIMBLE model
  Rmodel <- nimbleModel(code = code, constants = constants, data = data, inits = inits)
  Rmodel$calculate()

  # Build, compile, and run MCMC
  conf <- configureMCMC(Rmodel, monitors = c("mu[1]", "mu[2]", "sigma[1]", "sigma[2]", "theta", "lambda", "abundance", "z"))

  # Compile NIMBLE model into C for speed
  Rmcmc <- buildMCMC(conf, monitors = c("mu[1]", "mu[2]", "sigma[1]", "sigma[2]", "theta", "lambda", "abundance", "z"))
  Cmodel <- compileNimble(Rmodel, showCompilerOutput = TRUE)
  Cmcmc <- compileNimble(Rmcmc, project = Rmodel, showCompilerOutput = TRUE)


  set.seed(0) # Makes MCMC results replicable
  start_t <- Sys.time()
  continuous_results <- runMCMC(Cmcmc, niter = 3000, nburnin = 500, nchains = 2, samples = TRUE, summary = TRUE)
  end_t <- Sys.time()
  print(end_t - start_t)

  # Generate filename for results file
  results_filename <- sprintf("%sresults_%g.Rdata", results_save_path, real_lambda)

  save(continuous_results, file = results_filename)
}
```
```
